A highly-available move operation for replicated trees

### 2019-06-17 paper/move-op.tex
paper: make apply_op_commute2 inductive step more explicit

\noindent Let \isa{t1 = move\_time oper1} and \isa{t2 = move\_time oper2} be the timestamps of the two operations \isa{oper1} and \isa{oper2}.
Assume \isa{t1 < t2} without loss of generality.
We can now prove that lemma by induction over \isa{log}.

For the base case, where \isa{log} is empty, \isa{apply\_op oper2 (apply\_op oper1 (log, tree))} evaluates to first applying \isa{oper1} using \isa{do\_op}, then applying \isa{oper2} using \isa{do\_op}.
On the other hand, \isa{apply\_op oper1 (apply\_op oper2 (log, tree))} first applies \isa{oper2} using \isa{do\_op}, then undoes \isa{oper2} again to yield the original \isa{tree} (per \isa{do\_undo\_op\_inv}), then applies \isa{oper1} using \isa{do\_op}, and finally reapplies \isa{oper2} using \isa{redo\_op}.
The final states are the same on both sides of the equality.

For the inductive step, let the log be \isa{logop {\isacharhash} ops} (where \isa{\isacharhash} is the \emph{cons} operator), and let \isa{t3 = log\_time logop} be the timestamp of the operation at the head of the log.
We now consider three cases:
\begin{enumerate}
    \item \isa{t3 < t1 < t2}.
        Then neither applying \isa{oper1} nor applying \isa{oper2} will undo \isa{logop}.
        Thus, the two operations are applied in the same way as the base case.
    \item \isa{t1 < t3 < t2}.
        Then \isa{apply\_op oper2 (apply\_op oper1 (logop {\isacharhash} ops, tree))} first undoes \isa{logop}, then applies \isa{oper1} using \isa{do\_op}, then reapplies \isa{logop} using \isa{redo\_op}, and finally applies \isa{oper2} using \isa{do\_op}.
        On the other hand, \isa{apply\_op oper1 (apply\_op oper2 (logop {\isacharhash} ops, tree))} first applies \isa{oper2} using \isa{do\_op}, then undoes it again, then undoes \isa{logop}, applies \isa{oper1} and reapplies \isa{logop} as before, and finally reapplies \isa{oper2} using \isa{redo\_op}.
        The end result is the same.
    \item \isa{t1 < t2 < t3}.
        Then both applying \isa{oper1} and applying \isa{oper2} first undoes \isa{logop}, yielding \isa{tree' = undo\_op (logop, tree)}.
        From the inductive hypothesis we have
        \begin{quote}
        \isa{apply\_op oper2 (apply\_op oper1 (ops, tree')) =\\
             apply\_op oper1 (apply\_op oper2 (ops, tree'))}
        \end{quote}
        We now apply \isa{oper1} and \isa{oper2} as in the base case, and finally reapply \isa{logop} to both sides using \isa{redo\_op}, resulting in the same state on both sides.
\end{enumerate}

@@

>>>>

\noindent Let \isa{t1 = move\_time oper1} and \isa{t2 = move\_time oper2} be the timestamps of the two operations \isa{oper1} and \isa{oper2}.
Assume \isa{t1 < t2} without loss of generality.
We can now prove lemma \isa{apply\_op\_commute2} by induction over \isa{log}.

For the base case, where \isa{log} is empty, \isa{apply\_op oper2 (apply\_op oper1 ([], tree))} evaluates to first applying \isa{oper1} using \isa{do\_op}, then applying \isa{oper2} using \isa{do\_op}.
On the other hand, \isa{apply\_op oper1 (apply\_op oper2 ([], tree))} first applies \isa{oper2} using \isa{do\_op}, then undoes \isa{oper2} again to yield the original \isa{tree} (per \isa{do\_undo\_op\_inv}), then applies \isa{oper1} using \isa{do\_op}, and finally reapplies \isa{oper2} using \isa{redo\_op}.
The final states are the same on both sides of the equality.

For the inductive step, let the log be \isa{logop {\isacharhash} ops} (where \isa{\isacharhash} is the \emph{cons} operator).
We now need to show that
\begin{isabelle}
apply\_op oper2 (apply\_op oper1 (logop {\isacharhash} ops, tree)) =\\
apply\_op oper1 (apply\_op oper2 (logop {\isacharhash} ops, tree)).
\end{isabelle}
Let \isa{t3 = log\_time logop} be the timestamp of the operation at the head of the log.
We now consider three cases:
\begin{enumerate}
    \item \isa{t3 < t1 < t2}.
        Then neither applying \isa{oper1} nor applying \isa{oper2} will undo \isa{logop}.
        Thus, the two operations are applied in the same way as the base case.
    \item \isa{t1 < t3 < t2}.
        Then \isa{apply\_op oper2 (apply\_op oper1 (logop {\isacharhash} ops, tree))} first undoes \isa{logop}, then applies \isa{oper1} using \isa{do\_op}, then reapplies \isa{logop} using \isa{redo\_op}, and finally applies \isa{oper2} using \isa{do\_op}.
        On the other hand, \isa{apply\_op oper1 (apply\_op oper2 (logop {\isacharhash} ops, tree))} first applies \isa{oper2} using \isa{do\_op}, then undoes it again, then undoes \isa{logop}, applies \isa{oper1} and reapplies \isa{logop} as before, and finally reapplies \isa{oper2} using \isa{redo\_op}.
        The end result is the same.
    \item \isa{t1 < t2 < t3}.
        Then both applying \isa{oper1} and applying \isa{oper2} first undoes \isa{logop}, yielding \isa{tree' = undo\_op (logop, tree)}.
        From the inductive hypothesis we have
        \begin{quote}
        \isa{apply\_op oper2 (apply\_op oper1 (ops, tree')) =\\
             apply\_op oper1 (apply\_op oper2 (ops, tree'))}
        \end{quote}
        We now apply \isa{oper1} and \isa{oper2} as in the base case, and finally reapply \isa{logop} to both sides using \isa{redo\_op}, resulting in the same state on both sides.
\end{enumerate}

@@

\subsection{Performance experiments}

TODO\dots
